---
title: "Statistical inference with the GSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

```{r load-data}
load("gss.Rdata")
dim(gss)
```
There are total of 57061 observations and 114 variables in the GSS dataset.


```{r}
dfDegree <- gss %>% select(year, degree)
dfJobSat <- gss %>% select(year, satjob)
dfDegNjob <- gss %>% select(year, degree, satjob)
```


## Part 1: Data

The National Opinion Research Centre, University of Chicago has been collecting the data under the project entitled “General Social Survey (GSS) since the year 1972 onwards, a project funded by National Science Foundation. The aim of the GSS is to measure social change and trends and constants in attitudes, behaviour and attributes of the adult population of USA. Until 2012, information from about 57,000 respondents was collected.


METHODOLOGY

The target population of the GSS is adults (18+) living in households in the United States. The GSS sample is drawn using an area probability design that randomly selects respondents in households across the US to take part in the survey. The survey was conducted from 1972. A face to face survey is conducted by interviewing an adult from the selected house. He will be the single point of contact and further all the correspondence was made with him / her only. The survey takes about 90 minutes of time. 

SCOPE OF INFERENCE

The dataset consists of 57061 observations and 114 variables. The dataset is only an observational study. We can only make correlation study, associations based inferences. We cannot establish any causality from the dataset.





## Part 2: Research question

The aims of the project are to 

1. to study the effect of marital status on financial satisfaction.

2. to observer the relationship between Educational qualification and their Job Satisfaction.



## Part 3: Exploratory data analysis

Aim 1:

I am going to analyze the data and find out weather the financial satisfaction is connected to marital status. For this, I will be using the following variables.

satfin - financial situation
marital - Marital status

I am going to exclude the "NA" values from the data.


```{r}
gss_ma <- gss[ which(!is.na(gss$marital) & !is.na(gss$satfin)), ]
gss_ma <- gss_ma %>% select(marital, satfin)
dim(gss_ma)
```

There are 52441 observations in the data set that we are going to analyze excluding "NA".

The summary of the table is provided below.

```{r}
summary(gss_ma)
```


```{r}
table(gss_ma$marital, gss_ma$satfin)
```


```{r}
ggplot(aes(x=marital), data=gss_ma) + geom_bar(aes(fill=satfin), position = position_dodge())+ 
  ggtitle('Marital Status vs. Financial Satisfaction')
```

The bar graph shows that married people are more financially satisfied compared with that of others.







Aim 2:

I am interested to about the job satisfaction is related to education. 

The variables am going to use are

Educational Qualification - degree
job satisfaction - satjob



Firstly am going to study about the  Job Satisfaction trend and Academic Qualification Trend.




#### Educational Qualification trend

```{r}
dfDegFreq <- as.data.frame(prop.table(table(dfDegree), 1))
ggplot() + geom_bar(aes(y = Freq, x = year, fill = degree), data = dfDegFreq,
                    stat="identity") + coord_flip() +
    labs(x ="Year", y = "Relative Frequency" , fill = "Highest Degree")
```



####Job Satisfaction trend

```{r}
dfJobSatFreq <- as.data.frame(prop.table(table(dfJobSat), 1))
satColor <- c("Very Satisfied" = "#99FF00", "Mod. Satisfied" = "#CCFF33", "A Little Dissat" = "#FFFF00", "Very Dissatisfied" = "#CC9900")
ggplot() + geom_bar(aes(y = Freq, x = year, fill = satjob), data = dfJobSatFreq,
                    stat="identity") + coord_flip() + scale_fill_manual(values = satColor) +
    labs(x ="Year", y = "Relative Frequency" , fill = "Job Satisfaction")
```

####Educational Qualification and Job Satisfaction

```{r JobSatNdegree}
dfDegNjob <- dfDegNjob %>% filter(!is.na(degree), !is.na(satjob))
dfDegNjob$year <- as.character(dfDegNjob$year)
dfDegNjob$satIndex <- as.numeric(factor(dfDegNjob$satjob, levels = c("Very Dissatisfied", "A Little Dissat", "Mod. Satisfied", "Very Satisfied")))
dfDegNjob <- dfDegNjob %>% group_by(year, degree) %>% summarise(satIndexMean = mean(satIndex))
ggplot(dfDegNjob, aes(x = degree, y = year, fill = satIndexMean)) +
    geom_tile() +
    scale_fill_gradient(low = "red", high = "green", guide = guide_legend("Satisfaction Level"))
```


from the above three graphs, it is observed that education levels are increasing year by year and respondents are inclined more towards education. it is observed that job satisfaction is fluctuating because of various reasons. it is also observed that the graduates have more job satisfaction over the others.





## Part 4: Inference


Aim 1:

I am going to perform statistical inference on the results obtained. The process is indicated in the below steps:

1. Define hypothesis
2. Check for conditions
3. Perform the inference tests
4. Interpret the results



Define Hypothesis:

Null Hypothesis(H0H0): Marital status and satisfaction with financial situation are independent. Satisfaction with financial situation does not differ by marital status.

Alternative Hypothesis(HAHA): Marital status and satisfaction with financial situation are dependent. Satisfaction with financial situation does differ by marital status.


The defined hypothesis is as per our requirement to study if educational qualification and financial status are related to each other.



2. Check for conditions:

The dataset consists of two variables namely marital and satfin. I am going to use chi-square test of independence. It is the best test for comparing 2 categorical variables.

Sample size. As can be seen below, there are at least 5 counts for each cell

```{r}
gss_ma_table <- table(gss_ma$marital, gss_ma$satfin)
Xsq1 <- chisq.test(gss_ma_table)
Xsq1$expected
```

3. Perform the inference tests:

```{r}
(Xsq1 <- chisq.test(gss_ma_table))
```

We got the results that the Chi-Square statistic is 1728.2, degree of freedom is 8 and the associated P value is almost 0.


4. Interpret the results:

As the p value is very low and near to 0, reject the null hypothesis. Hence it is conclude that satisfaction with financial situation and marital status are dependent. Since the study is only a observational, we can only estimate association. we cannot estimate the casualties between the the two variables.




Aim 2:

I am going to perform statistical inference on the results obtained. The process is indicated in the below steps:

1. Define hypothesis
2. Check for conditions
3. Perform the inference tests
4. Interpret the results



Define hypothesis:

The null hypothesis (H0H0): It is to show that the highest degree obtained and their job satisfaction level are independent.

The alternative hypothesis (HAHA): It is to show that the highest degree obtained and their job satisfaction level are not independent.


```{r variables}
dfDegNjob <- gss %>% select(year, degree, satjob)
str(dfDegNjob)
```

2. Check for conditions


The dataset consists of two variables namely degree and satjob. I am going to use chi-square test of independence. It is the best test for comparing 2 categorical variables.

Sample size. As can be seen below, there are at least 5 counts for each cell.


```{r sampleSize}
table(dfDegNjob$degree,dfDegNjob$satjob)
```


3. Perform the inference tests:

With all the conditions met, we will now work on the inference calculation using the chi-square test.

```{r}
chisq.test(dfDegNjob$degree,dfDegNjob$satjob)
```


We got the results that the Chi-Square statistic is 267.14, degree of freedom is 12 and the associated P value is almost 0.


4. Interpret the results:

As the p value is very low and near to 0, reject the null hypothesis. Therefore it is conclude that there is evidence that the educational qualification and job satisfaction are related.. Since the study is only a observational, we can only estimate association. we can estimate the casualties between the the two variables.





